---
title: "Aves Synonymy"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

This is an R Markdown for the TNC Gap Analysis to complete the synonymy for Aves. DAK has done the Cal-IBIS and SBMNH synonymy. Using lists created from her completed synonymy this code will complete the synonymy for the iNaturalist and eBird datasets. 


#Download Packages & Data
```{r Packages!}
library(tidyverse)
library(lubridate)
library(stringr)
```

```{r Data}
aves_syn = read_csv("Aves/aves_syn.csv")
aves_inat_messy = read_csv("Aves/aves_inat_messy.csv")
aves_ebird_messy = read_csv("Aves/aves_ebird_messy.csv")

```

```{r}
aves_inat1 <- left_join(aves_inat_messy, aves_syn)
write.csv(aves_inat1, "Aves/aves_inat1.csv")
#about 1000 lines missing, 162 unique ScientificNames

aves_ebird1 <- left_join(aves_ebird_messy, aves_syn)
write.csv(aves_ebird1, "Aves/aves_ebird1.csv")
#about 2000 lines missing, 34 unique ScientificNames (179 inat and ebird combined)


```

Denise did the synonymy for the missing lines, and I have an updated list to compare the inat and ebird data too, will repeat the same process as above with the new list and the "aves_inat1" & "aves_ebird1" files 

```{r}
aves_syn2 = read_csv("Aves/aves_syn2.csv")
aves_syn_complete = rbind(aves_syn, aves_syn2)

aves_inat2 <- left_join(aves_inat_messy, aves_syn_complete)
aves_ebird2 <- left_join(aves_ebird_messy, aves_syn_complete)

write.csv(aves_inat2, "Aves/aves_inat2.csv")
write.csv(aves_ebird2, "Aves/aves_ebird2.csv")
#have to do some manual corrects on inat2 due to funky characters 
```

