---
title: "Gap Analysis Hymenoptera Combined datasets"
author: "Kylie Etter"
date: "7/8/2021"
output: pdf_document
editor_options: 
  chunk_output_type: console
---
This R Code is for data analysis of the Hymenoptera for the TNC Gap Analysis. It downloads the cleaned datasets to combine that have (source, month_arabic, day_arabic, year, island, order, bee, family, genera, specificEp, scientificName) columns. Once combined can compare numbers between sources, compare locality and temporal information. 

The iNaturalist data was removed from the Cal-IBIS data so there was no duplication. 

# Download datasets. (iNaturalist, LACM, Cal-IBIS, Bohart)
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load packages
library(tidyverse)
library(lubridate)


inat =read.csv("~/gitit/TNC Gap Analysis/hym_inat_clean.csv")
lacm = read.csv("~/gitit/TNC Gap Analysis/hym_lacm_clean.csv")
chis = read_csv("~/gitit/TNC Gap Analysis/hym_chis_clean.csv")
bohart = read.csv("~/gitit/TNC Gap Analysis/hym_bohart_clean.csv")

```

#Clean up datasets and combine 

Combine datasets with the specific columns (source, month_arabic, day_arabic, year, island, order, family, genus, specificEp, scientificName). 

First, removing excess columns from datasets.

```{r clean up datasets}
inat_column <- inat %>%
  select(source, month_arabic, day_arabic, year, island, order, family, genus, specificEp, scientificName)
lacm_column <- lacm %>%
  select(source, month_arabic, day_arabic, year, island, order, family, genus, specificEp, scientificName) 
chis_column <- chis %>%
  select(source, month_arabic, day_arabic, year, island, order, family, genus, specificEp, scientificName) %>%
filter(!is.na(source))
bohart_column <- bohart %>%
  select(source, month_arabic, day_arabic, year, island, order, family, genus, specificEp, scientificName)

```

Make master dataset.
Now to combine. iNat = 1372, LACM = 1186, CHIS = 7621, Bohart = 6326 (bohart is all bees). 16505 total. 

```{r create the master }
masterdata = rbind(inat_column, lacm_column, chis_column, bohart_column)
write.csv(masterdata, "hym_master.csv")
```

Updated master dataset with the synonymy. 
```{r}
mastersyn = read.csv("~/gitit/TNC Gap Analysis/hym_master_syn.csv")

```


#Quick and dirty summary stats
Now all 4 data sets are combine! Summary stats here we go! 

```{r master summary}
x= table(masterdata$order)
x
table(masterdata$family)
sum(is.na(masterdata$family))
sum(is.na(masterdata$genus))
sum(is.na(masterdata$specificEp))

colSums(is.na(masterdata))

```
Summary of taxonomic resolution by source 
```{r summary by data source}
table(inat_column$order)
table(inat_column$family)
x = table(inat_column$family)
print(x)

table(chis_column$family)

table(bohart_column$family)

table(lacm_column$family)

table(masterdata$family)


write.csv(masterdata, "gapan_alldata.csv")

```

#Graphs of the different data. 

Mapping the sampling on different islands by data source.
```{r Graphing by island}
library(ggplot2)
 
y= ggplot(masterdata, aes(x=island, fill=source)) + geom_bar(position="fill")
y

r = ggplot(masterdata, aes(x=year, fill=source)) + geom_bar(position="fill")
r

masterdataisland <- masterdata %>% filter(!is.na(island))
masterdataisland <- masterdataisland[c(1:1371, 1373:16366),]

islandplot = ggplot(data = masterdataisland, aes(x = island, fill=source)) + geom_bar()
islandplot = islandplot + theme_classic() + labs(x="Island", y= "Number of Specimen Collected")
qw = islandplot +  stat_count(geom = "text", 
             aes(label = stat(count)),
             position=position_fill(vjust=1), colour="black")

islandplot2 = ggplot(data = masterdataisland, aes(x = island, fill = source)) +
    geom_bar(position = "dodge")
islandplot2 = islandplot2 + theme_classic() + labs(x="Island", y= "Number of Specimen Collected")
wq = islandplot2 + stat_count(geom = "text", 
             aes(label = stat(count)),
             position= "dodge", colour="black")

islandall = ggplot(data = masterdataisland, aes(x=island)) + geom_bar()
islandall = islandall + stat_count(geom = "text", 
             aes(label = stat(count)),
             position=position_fill(vjust=0.5), colour="black")
islandall + theme_classic()



```
# With the Synonymy done 
```{r}

mastersynisland <- mastersyn %>% filter(!is.na(island))
mastersynisland <- mastersynisland[c(1:1371, 1373:16366),]

islandplot = ggplot(data = mastersynisland, aes(x = island, fill=source)) + geom_bar()
islandplot = islandplot + theme_classic() + labs(x="Island", y= "Number of Specimen Collected")
qw = islandplot +  stat_count(geom = "text", 
             aes(label = stat(count)),
             position=position_fill(vjust=1), colour="black")

islandplot2 = ggplot(data = mastersynisland, aes(x = island, fill = source)) +
    geom_bar(position = "dodge")
islandplot2 = islandplot2 + theme_classic() + labs(x="Island", y= "Number of Specimen Collected")
wq = islandplot2 + stat_count(geom = "text", 
             aes(label = stat(count)),
             position= "dodge", colour="black")

islandall = ggplot(data = mastersynisland, aes(x=island)) + geom_bar()
islandall = islandall + stat_count(geom = "text", 
             aes(label = stat(count)),
             position=position_fill(vjust=0.5), colour="black")
islandall + theme_classic()


```


Graphing the temporal gaps in collections 
```{r Graphing by year}
yearplot = ggplot(masterdata, aes(x=year, fill=source)) + geom_bar()
yearplot = yearplot + theme_classic() + labs(x="Year", y ="Number of Specimen Collected")
yearplot

yearplot2 = ggplot(masterdata, aes(x=year, fill=source)) + geom_bar(position="dodge")
yearplot2 = yearplot2 + theme_classic() + labs(x="Year", y ="Number of Specimen Collected")
yearplot2
```

Graphing the seasonal gaps in collections. 

```{r Graphing by month}
monthplot = ggplot(masterdata, aes(x=month_arabic, fill=source)) + geom_bar() + scale_x_continuous(breaks = seq(1, 12, by=1)) 
monthplot = monthplot + theme_classic() + (labs(x="Month", y="Number of Specimen Collected")) + scale_x_continuous(breaks = seq(1,12, by=1))
monthplot

monthplotall = ggplot(masterdata, aes(x=month_arabic)) + geom_bar()
monthplotall = monthplotall + theme_classic() + (labs(x="Month", y="Number of Specimen Collected")) + scale_x_continuous(breaks = seq(1, 12, by=1)) 
zx = monthplotall + stat_count(geom = "text", 
             aes(label = stat(count)),
             position=position_fill(vjust=0.5), colour="black")

```

# Unique values by datasource 
```{r}
library(plyr)
ddply(mastersyn,~source,summarize,number_of_distint_fam=length(unique(family)))

ddply(mastersyn,~source,summarize,number_of_disitinct_gen=length(unique(genus)))

ddply(mastersyn,~source,summarize,number_of_disitinct_gen=length(unique(acceptedScientificName)))

chis_fam = unique(c(as.character(chis$family)))
unique_family =as.data.frame(chis_fam)
bohart_fam = unique(c(as.character(bohart$family)))
bohart_fam


cbind(mastersyn, record)
```

 
# Playing around with pie graphs!!!
```{r}
ggplot(masterdata ,aes(x = "", fill=genus)) + geom_bar(width =
1) + coord_polar (theta="y")
```

