---
title: "Invert_Endemics"
author: "Kylie Etter"
date: "9/27/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

This is a R Markdown to make graphs for the Channel Islands endemic invertebrates for the TNC Gap Analysis. The master occurence data files were made in "Invert_Code.Rmd" from the various sources per taxa and then saved to the project. I am using the masster master data file created for this code. To do:aran, C3,C4,C8
C5
C7
C9
dipt
hemi
bee
wasp
ant
diplo
L1
L4
butt&skip
L6
L7
opil
grass
crick
phas
scor
moll1moll2,moll3,moll4


#Download data
```{r Endemic List}
invert_end <- read_csv("invert_endemic2.csv") %>%
  rename(endemic_source = "source", islands="Island(s)")
```
```{r Master Taxa Data}
master_inverts <-read_csv("master/master_inverts2.csv")
```

#Join endemic data to master data
```{r join data}
invert_end_join = inner_join(master_inverts, invert_end, by="acceptedName")

#Looking at what is on the endemic list but not in our data 
invertdata = unique(c(as.character(invert_end_join$acceptedName)))
invertend = unique(c(as.character(invert_end$acceptedName)))
invertdiff = setdiff(invertend, invertdata)
invertdiff = as.data.frame(invertdiff) %>% 
  rename(acceptedName="invertdiff")
invertdiff2 = left_join(invertdiff, invert_end)

#81 not in our data....

#updating synonymy for invert_end with synonymy list created from masters, and creating csv of result to go through unsyn and eliminate or write in synonymy that I will then use to update the synonymy list / confirm really not in our data
master_invert_syn = read_Csv("master/master_invert_syn.csv")
invert_end = invert_end %>% rename(scientificName = "acceptedName")
invert_end_syn = left_join(invert_end, master_invert_syn)
write_csv(invert_end_syn, "invert_end_syn.csv", na="")
```

```{r}
output <- matrix(data=invert_end$islands, nrow=151, ncol = 8, byrow = FALSE)
island = as.data.frame(output)
endemics =cbind(invert_end, island)
write_csv(endemics, "endemics/invert_endemic_list1.csv", na="")
```
```{r}
invert_end_list = read_csv("endemics/invert_endemic_list1_clean.csv")


invertcount_occ = table(invert_end_join$acceptedName)
invertcount_occ = as.data.frame.table(invertcount_occ) %>%
  rename(Occurence= "Freq",  acceptedName = "Var1")

invert_end_list2 = left_join(invert_end_list, invertcount_occ)
write_csv(invert_end_list2, "endemics/invert_endemic_list2.csv", na="")
```

#Arachnida 
```{r Acari}
arac_end = filter(invert_end_join, group_forR=="aran")
arac_islandsum <- arac_end %>%
  mutate(no_sample = 1) %>%
  group_by(island, acceptedName, source) %>%
  summarise(samples = sum(no_sample))

a_m_islandsum2 <- na.omit(subset(aves_m_islandsum, select = c(island, samples, source2, acceptedName)))

island_size = read_csv("island_size.csv")
arac_islandsum2 = merge(arac_islandsum, island_size, by="island", all.y=TRUE) 

x= c("#D95F02")

#replaced NAs in acceptedName with one of the species so an NA graph doesn't appear, but doesn't impact other graphs at all
arac_islandsum2$acceptedName[is.na(arac_islandsum2$acceptedName)] <- "Lutica nicolasia"

arac_island <- ggplot(data = arac_islandsum2, aes(x = samples, y = fct_rev(island), fill=source)) + geom_bar(stat = "identity") + facet_wrap(vars(acceptedName), ncol=2, scales="free_x") +
  labs(y ="Island",  x = "Number of Data Points", fill = "Data Source") + theme_bw() + scale_fill_manual(values=x) + scale_x_continuous(expand=expansion(mult=c(0,0.05))) + theme(axis.text = element_text(size = 8), axis.title = element_text(size=15), strip.text = element_text(size=14), legend.text = element_text(size=12)) 

arac_island

ggsave(plot=arac_island, filename="endemics/InvertGraphs/arac_island.png", height=10, width=35, units="cm", dpi=300)
```

#Coleoptera
##C5
```{r}
#brewer.pal(5, name="Dark2") Color codes for the brewer colors, need 1, 2 and 4 for this one
#Cal-IBIS "#1B9E77"            iNaturalist "#D95F02"            EBIRD #7570B3"
#SBMNH Col "#E7298A"            SBMNH Obs "#7FBC41"

col5_end = filter(invert_end_join, group_forR=="C5")
col5_islandsum <- col5_end %>%
  mutate(no_sample = 1) %>%
  group_by(island, acceptedName, source) %>%
  summarise(samples = sum(no_sample))



island_size = read_csv("island_size.csv")
col5_islandsum2 = merge(col5_islandsum, island_size, by="island", all.y=TRUE) 

x= c("#1B9E77", "#7570B3")

#replaced NAs in acceptedName with one of the species so an NA graph doesn't appear, but doesn't impact other graphs at all
col5_islandsum2$acceptedName[is.na(col5_islandsum2$acceptedName)] <- "Serica cruzi"

col5_island <- ggplot(data = col5_islandsum2, aes(x = samples, y = fct_rev(island), fill=source)) + geom_bar(stat = "identity") + facet_wrap(vars(acceptedName), ncol=2, scales="free_x") +
  labs(y ="Island",  x = "Number of Data Points", fill = "Data Source") + theme_bw() + scale_fill_manual(values=x) + scale_x_continuous(expand=expansion(mult=c(0,0.05))) + theme(axis.text = element_text(size = 8), axis.title = element_text(size=15), strip.text = element_text(size=14), legend.text = element_text(size=12)) 

col5_island

ggsave(plot=col5_island, filename="endemics/InvertGraphs/col5_island.png", height=20, width=35, units="cm", dpi=300)
```
#C7
```{r}
#brewer.pal(5, name="Dark2") Color codes for the brewer colors, need 1, 2 and 4 for this one
#Cal-IBIS "#1B9E77"            iNaturalist "#D95F02"            EBIRD #7570B3"
#SBMNH Col "#E7298A"            SBMNH Obs "#7FBC41"

col7_end = filter(invert_end_join, group_forR=="C7")
col7_islandsum <- col7_end %>%
  mutate(no_sample = 1) %>%
  group_by(island, acceptedName, source) %>%
  summarise(samples = sum(no_sample))



island_size = read_csv("island_size.csv")
col7_islandsum2 = merge(col7_islandsum, island_size, by="island", all.y=TRUE) 

x= c("#1B9E77", "#D95F02", "#7570B3")

#replaced NAs in acceptedName with one of the species so an NA graph doesn't appear, but doesn't impact other graphs at all
col7_islandsum2$acceptedName[is.na(col7_islandsum2$acceptedName)] <- "Coenonycha fulva"

col7_island <- ggplot(data = col7_islandsum2, aes(x = samples, y = fct_rev(island), fill=source)) + geom_bar(stat = "identity") + facet_wrap(vars(acceptedName), ncol=2, scales="free_x") +
  labs(y ="Island",  x = "Number of Data Points", fill = "Data Source") + theme_bw() + scale_fill_manual(values=x) + scale_x_continuous(expand=expansion(mult=c(0,0.05))) + theme(axis.text = element_text(size = 8), axis.title = element_text(size=15), strip.text = element_text(size=14), legend.text = element_text(size=12)) 

col7_island

ggsave(plot=col7_island, filename="endemics/InvertGraphs/col7_island.png", height=20, width=35, units="cm", dpi=300)
```


