---
title: "Invert Code"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

This is an R Markdown for the Gap Analysis Invertebrate Graph and Table Making. This is the cleaned up code from my first run with Hymenoptera, this code should be able to be used with all the datasets for each Order with little adjustment. 

# Diptera 

## Dip Step 1. Download Packages and Import Data
```{r setup}
knitr::opts_chunk$set(echo = TRUE)

# load packages
library(tidyverse)
library(lubridate)
library(ggplot2)
library(dplyr)
library(RColorBrewer)

dip_inat <- read_csv("diptera/dip_inat.csv")
dip_lacm <- read_csv("diptera/dip_lacm.csv")
dip_chis <- read_csv("diptera/dip_chis.csv")
dip_bold <- read_csv("diptera/dip_bold.csv")

# extra for flies
dip_nic <- read_csv("diptera/dip_nic.csv")

```


## Dip Step 2. Combine datasets 
```{r}
dip_inat_combine <- dip_inat %>%
  select(source, month, year, island, order, acceptedFamily, acceptedGenus, acceptedName)

dip_lacm_combine <- dip_lacm %>%
   select(source, month, year, island, order, acceptedFamily, acceptedGenus, acceptedName)

dip_chis_combine <- dip_chis %>%
  select(source, month, year, island, order, acceptedFamily, acceptedGenus, acceptedName)

dip_nic_combine <- dip_nic %>%
  select(source, month, year, island, order, acceptedFamily, acceptedGenus, acceptedName)

dip_master = rbind(dip_inat_combine, dip_chis_combine, dip_lacm_combine, dip_nic_combine) %>%
  # reorder the island names to show up N -> S on list
  mutate(island = fct_relevel(island, "San Miguel", "Santa Rosa", "Santa Cruz", "Anacapa", "Santa Barbara", "Santa Catalina", "San Nicolas", "San Clemente"))

# bold doesn't have date info
dip_bold_combine <-dip_bold %>%
   select(source, island, acceptedFamily, acceptedGenus, acceptedName) 

dip_combine <- dip_master %>%
   select(source, island, acceptedFamily, acceptedGenus, acceptedName) 

dip_master_bold = rbind(dip_combine, dip_bold_combine)
```

```{r heat map}
dip_inat_map <- dip_inat %>%
  select(source, month, year, island, order, acceptedFamily, acceptedGenus, acceptedName, latitude, longitude)

```

##Dip Step. 3 Make Graphs! 

### Island Graph 
```{r Graphing by island}

dip_islandsum <- dip_master %>%
  mutate(no_sample = 1) %>%
  group_by(island, source) %>%
  summarise(samples = sum(no_sample))

dip_island <- ggplot(data = dip_islandsum, aes(x = samples, y = fct_rev(island), fill = source)) + 
  geom_bar(stat = "identity") +
  labs(x ="Island", 
       y = "Number of Specimens Collected",
       fill = "Data Source") +
  theme_bw() + 
  scale_x_continuous(limits =c(0,1200), expand = c(0,0)) +
  scale_fill_brewer(palette = "Dark2")

dip_island 

ggsave(plot=dip_island, filename="diptera/DipteraGraphs/dip_island.png", height=5, width=8, units="in", dpi=300)

```

```{r Graphing by island proportional}
island_size = read_csv("island_size.csv")
dip_islandsum2 = merge(dip_islandsum, island_size, by="island", all.y=TRUE)
dip_islandprop = dip_islandsum2 %>% group_by(island) %>% mutate(by_area = samples/island_size_85)

dip_island_prop <- ggplot(data = dip_islandprop, aes(x = by_area, y = fct_rev(island), fill = source)) + 
  geom_bar(stat = "identity") +
  labs(x ="Island", 
       y = "Specimens Collected/Island Area (km^2)",
       fill = "Data Source") +
  theme_bw() + 
  scale_x_continuous(limits =c(0,70), expand = c(0,0)) +
  scale_fill_brewer(palette = "Dark2")

dip_island_prop 

ggsave(plot=dip_island_prop, filename="diptera/DipteraGraphs/dip_island_prop.png", height=5, width=8, units="in", dpi=300)
```


### Temporal Graph
```{r Graphing by year}


dip_year = ggplot(dip_master, aes(x=year, fill=source)) + 
  geom_bar() + 
  theme_bw() + 
  scale_y_continuous(expand=c(0,0), limits = c(0, 1000)) + 
  scale_x_continuous(limits=c(1910, 2025)) + 
  labs(x="Year", y ="Number of Specimens Sampled") +   scale_fill_brewer(palette = "Dark2")

dip_year

ggsave(plot=dip_year, filename="diptera/DipteraGraphs/dip_year.png", height=5, width=9, units="in", dpi=300)

```

###Graphing the seasonal gaps in collections. 

```{r Graphing by month}

occmonth <- ggplot(mastersyn, aes(x=month, fill=source)) + 
  geom_bar() + 
  theme_bw() + 
  scale_y_continuous(expand=c(0,0)) + 
  labs(x="Month", y ="Number of Specimens Samples") +
  scale_x_continuous(breaks = seq(1, 12, by=1)) 

occmonth

# by month and island 
dip_master$month <- as.factor(dip_master$month)
dip_month <- na.omit(subset(dip_master, select = c(island, month)))


dip_ismonth <- ggplot(dip_month, aes(x=month, fill=month)) + 
  geom_bar() + 
  theme_bw() + 
  scale_y_continuous(expand=c(0,0.5)) + 
  labs(x="Sampling per Island by Month", y ="Number of Specimens Sampled") +  
  facet_wrap(vars(island), nrow=2, scales="free_y")

dip_ismonth 

ggsave(plot=dip_ismonth, filename="diptera/DipteraGraphs/dip_seasonal.png", height=5, width=9, units="in", dpi=300)

```

##Dip Step 4. Make Tables 
```{r Family by Source Table}
q = table(dip_master_bold$acceptedFamily)
q = as.data.frame.table(q)
write.csv(q, "diptera/DipteraGraphs/dip_table.csv")

# include BOLD? so few 
table(dip_bold_combine$acceptedFamily)

table(dip_chis_combine$acceptedFamily)

table(dip_lacm_combine$acceptedFamily)

table(dip_inat_combine$acceptedFamily)

table(dip_nic_combine$acceptedFamily)


```

```{r Source Resolution}
ddply(dip_master_bold, "source", summarise, n = length(unique(acceptedFamily)))

ddply(dip_master_bold, "source", summarise, n = length(acceptedFamily))
ddply(dip_master_bold, "source", summarise, n = length(acceptedGenus))
ddply(dip_master_bold, "source", summarise, n = length(acceptedName))

```

#Mammals

##Mam Step 1. Download Packages and Import Data
```{r setup}
knitr::opts_chunk$set(echo = TRUE)

# load packages
library(tidyverse)
library(lubridate)
library(ggplot2)
library(dplyr)
library(RColorBrewer)

mam_inat <- read_csv("mammal/mam_inat.csv")

mam_chis <- read_csv("mammal/mam_chis.csv")

```


##Mam Step 2. Combine datasets 
```{r}
#inat doesn't have order listed..would need to add 
mam_inat_combine <- mam_inat %>%
  rename(class="iconic_tax") %>%
  select(source, month, year, island, class, acceptedFamily, acceptedGenus, acceptedName)

mam_chis_combine <- mam_chis %>%
  select(source, month, year, island, class, acceptedFamily, acceptedGenus, acceptedName)


mam_master = rbind(mam_inat_combine, mam_chis_combine) %>%
  # reorder the island names to show up N -> S on list
  mutate(island = fct_relevel(island, "San Miguel", "Santa Rosa", "Santa Cruz", "Anacapa", "Santa Barbara", "Santa Catalina", "San Nicolas", "San Clemente"))

```

```{r heat map}
mam_inat_map <- mam_inat %>%
  select(source, month, year, island, class, acceptedFamily, acceptedGenus, acceptedName, latitude, longitude)

mam_chis_map <- mam_inat %>%
  rename(latitude = "decimalLatitude", longitude="decimalLongitude") %>%
  select(source, month, year, island, class, acceptedFamily, acceptedGenus, acceptedName, latitude, longitude)

mam_map = rbind(mam_inat_map, mam_chis_map)

```

##Mam Step. 3 Make Graphs! 

### Island Graph 
```{r Graphing by island}

mam_islandsum <- mam_master %>%
  mutate(no_sample = 1) %>%
  group_by(island, source) %>%
  summarise(samples = sum(no_sample))
mam_islandsum <- na.omit(subset(mam_islandsum, select = c(island, samples, source)))

mam_island <- ggplot(data = mam_islandsum, aes(x = samples, y = fct_rev(island), fill = source)) + 
  geom_bar(stat = "identity") +
  labs(x ="Island", 
       y = "Number of Mammal Specimens Collected",
       fill = "Data Source") +
  theme_bw() + 
  scale_x_continuous(limits =c(0,1200), expand = c(0,0)) +
  scale_fill_brewer(palette = "Dark2")

mam_island 

ggsave(plot=mam_island, filename="mammal/MammalGraphs/mam_island.png", height=5, width=8, units="in", dpi=300)

```

```{r Graphing by island proportional}
island_size = read_csv("island_size.csv")
mam_islandsum2 = merge(mam_islandsum, island_size, by="island", all.y=TRUE)
mam_islandprop = mam_islandsum2 %>% group_by(island) %>% mutate(by_area = samples/island_size_85)

mam_island_prop <- ggplot(data = mam_islandprop, aes(x = by_area, y = fct_rev(island), fill = source)) + 
  geom_bar(stat = "identity") +
  labs(x ="Island", 
       y = "Mammal Specimens Collected/Island Area (km^2)",
       fill = "Data Source") +
  theme_bw() + 
  scale_x_continuous(limits =c(0,10), expand = c(0,0)) +
  scale_fill_brewer(palette = "Dark2")

mam_island_prop 

ggsave(plot=mam_island_prop, filename="mammal/MammalGraphs/mam_island_prop.png", height=5, width=8, units="in", dpi=300)
```


### Temporal Graph
```{r Graphing by year}


mam_year = ggplot(mam_master, aes(x=year, fill=source)) + 
  geom_bar() + 
  theme_bw() + 
  scale_y_continuous(expand=c(0,0), limits = c(0, 900)) + 
  scale_x_continuous(expand=c(0,0), limits=c(1885, 2020), breaks=c(1885, 1900, 1915, 1930, 1945, 1960, 1975, 1990, 2005, 2020)) + 
  labs(x="Year", y ="Number of Mammal Specimens Sampled") +   scale_fill_brewer(palette = "Dark2")

mam_year

ggsave(plot=mam_year, filename="mammal/MammalGraphs/mam_year.png", height=5, width=9, units="in", dpi=300)

```

###Graphing the seasonal gaps in collections. 

```{r Graphing by month}
# by month and island 
mam_master$month <- as.factor(mam_master$month)
mam_month <- na.omit(subset(mam_master, select = c(island, month)))


mam_ismonth <- ggplot(mam_month, aes(x=month, fill=month)) + 
  geom_bar() + 
  theme_bw() + 
  scale_y_continuous(expand=c(0,0.5)) + 
  labs(x="Mammal Sampling per Island by Month", y ="Number of Specimens Sampled") +  
  facet_wrap(vars(island), nrow=2, scales="free_y")

mam_ismonth 

ggsave(plot=mam_ismonth, filename="mammal/MammalGraphs/mam_seasonal.png", height=5, width=9, units="in", dpi=300)

```

```{r Mammal Collection Type}
mam_coll <- ggplot(mam_chis, aes(x=CollectionType, fill=CollectionType)) + geom_bar() + theme_bw() + theme(legend.position = "none") + theme(axis.text.x = element_text(angle = 90))

mam_coll
ggsave(plot=mam_coll, filename="mammal/MammalGraphs/mam_type.png", height=5, width=8, units="in", dpi=300)
```


##Mam Step 4. Make Tables 
```{r Family by Source Table}
mamtab = table(dip_master_bold$acceptedFamily)
mamtab = as.data.frame.table(mamtab)
write.csv(mamtab, "mammal/MammalGraphs/mam_table.csv")


table(dip_chis_combine$acceptedFamily)


table(dip_inat_combine$acceptedFamily)



```

```{r Source Resolution}
ddply(dip_master_bold, "source", summarise, n = length(unique(acceptedFamily)))

ddply(dip_master_bold, "source", summarise, n = length(acceptedFamily))
ddply(dip_master_bold, "source", summarise, n = length(acceptedGenus))
ddply(dip_master_bold, "source", summarise, n = length(acceptedName))

```